The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-06-07 Fsing/FThresholdsPaper/FThresholds.tex
Try to fix crlf
Added a paragraph.
Fixed a dumb issue.
Replaced Section 3 with updated version

i21 : fpt(f, DepthOfSearch => 3, Attempts => 4) -- and one more finds

      the answer

       1
o21 = --
      19

o21 : QQ
\end{MyVerbatim}
}
\medspace

If {\tt guessFPT} is unsuccessful and {\tt UseFSignature} is set to {\tt true}, then the {\tt fpt} function proceeds to use the convexity of the $F$-signature function and a secant line argument to attempt to narrow down the interval bounding the $F$-pure threshold.

@@

i22 : f = x^5*y^6*(x + y)^9*(x^2 + y^3)^4;

i23 : fpt(f, DepthOfSearch => 3)

        2   1
o23 = {--, --}
       31  14

o23 : List

i24 : fpt(f, DepthOfSearch => 3, UseFSignature => true)

        181   1
o24 = {----, --}
       2750  14

@@

i25 : numeric o23

o25 = {.0645161, .0714286}

o25 : List

i26 : numeric o24 -- UseFSignature sharpened the estimate a bit

o26 = {.0658182, .0714286}

o26 : List
\end{MyVerbatim}
}
\medspace

When {\tt FRegularityCheck} is set to {\tt true} and no exact answer has been found, a final check is run to verify whether the final lower bound for the $F$-pure threshold is the exact answer (if it has not already been eliminated as a possibility).

@@

i27 : f = (x + y)^4*(x^2 + y^3)^6;

i28 : fpt( f, Attempts => 2, DepthOfSearch => 3 )

        3   13
o28 = {--, ---}
       31  125

o28 : List

i29 : fpt(f, Attempts => 2, DepthOfSearch => 3, UseFSignature => true)

      -- UseFSignature improves the answer a bit

        1   13
o29 = {--, ---}
       10  125

o29 : List

i30 : fpt(f, Attempts => 2, DepthOfSearch => 3, UseFSignature => true,

      FRegularityCheck => true) -- FRegularityCheck finds the answer

       1
o30 = --
      10

o30 : QQ
\end{MyVerbatim}
}
\medspace

The computations performed when {\tt UseFSignature} and {\tt FRegularityCheck} are set to {\tt true} are often slow, and often fail to improve the estimate, and for this reason, these options should be used sparingly.
            It is often more effective to increase the values of {\tt Attempts} or {\tt DepthOfSearch} instead.

@@

i31 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

i32 : timing numeric fpt(f, DepthOfSearch => 3, UseFSignature => true,

      FRegularityCheck => true)

o32 = {.0733061, .0769231}
      -- 2.68899 seconds

o32 : Time

i33 : timing numeric fpt(f, Attempts => 5, DepthOfSearch => 3) -- a better

      answer in less time

o33 = {.075, .0769231}
      -- .893389 seconds

o33 : Time

i34 : timing fpt(f, DepthOfSearch => 4) -- the exact answer in even less
      time

       48
o34 = ---
      625
      -- .361882 seconds

o34 : Time
\end{MyVerbatim}
}
\medspace

As seen in several examples above, when the exact  $F$-pure threshold of $f$ is not found, a list containing the endpoints of an interval containing its value is returned.  Whether that interval is open, closed, or a mixed interval depends on the options passed; if the option {\tt Verbose} is set to {\tt true}, the precise interval will be printed.

i35 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

i36 : fpt(f, DepthOfSearch => 3, UseFSignature => true, Verbose => true)

@@

guessFPT narrowed the interval down to ( 9/124, 1/13 ) ...

@@

F-signature intercept is an improved lower bound ...

fpt failed to find the exact answer; try increasing the value of

DepthOfSearch or Attempts.

fpt lies in the interval [ 449/6125, 1/13 ).

        449   1
o36 = {----, --}
       6125  13

o36 : List
\end{MyVerbatim}
}
\medspace

>>>>

i21 : fpt(f, DepthOfSearch => 3, Attempts => 4) -- more attempts *always* 
      improve the estimate

        1   4
o21 = {--, --}
       20  75

o21 : List

i22 : fpt(f, DepthOfSearch => 3, Attempts => 4, GuessStrategy => denominator) 
      -- the exact answer, after a change of strategy

       1
o22 = --
      19

o22 : QQ
\end{MyVerbatim}
}
\medspace

The option {\tt Bounds} allows the user to specify known lower and upper bounds for the $F$-pure threshold of $f$, in order to speed up computations or to refine previously obtained estimates.

@@

i23 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

i24 : fpt(f, DepthOfSearch => 3, Attempts => 5)

        3   1
o24 = {--, --}
       40  13

@@

i25 : fpt(f, DepthOfSearch => 3, Attempts => 5, Bounds => oo)

        24   1
o25 = {---, --}
       313  13

o25 : List
\end{MyVerbatim}
}
\medspace

% f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;
% fpt(f, DepthOfSearch => 3, Attempts => 5)
% fpt(f, DepthOfSearch => 3, Attempts => 5, Bounds => oo)

If {\tt guessFPT} is unsuccessful and {\tt FinalAttempt} is set to {\tt true}, the {\tt fpt} function proceeds to use the convexity of the $F$-signature function and a secant line argument to attempt to narrow down the interval bounding the $F$-pure threshold.
If successful, the new lower bound may coincide with the upper bound, in which case we can conclude that it is the desired $F$-pure threshold.
If this is not the case, an $F$-regularity check is done at the new lower bound, to verify if it is the $F$-pure threshold.

@@

i26 : f = -2*x^10*y^5 - x^5*y^9 - 2*x^3*y^10 + 2*x^2*y^8 - 2*x*y^9;

i27 : numeric fpt(f, DepthOfSearch => 3)

o27 = {.153226, .158333}

o27 : List

i28 : numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
      -- FinalAttempt improves the estimate slightly

o28 = {.153461, .158333}

o28 : List
\end{MyVerbatim}
}
\medspace

% f = -2*x^10*y^5 - x^5*y^9 - 2*x^3*y^10 + 2*x^2*y^8 - 2*x*y^9;
% numeric fpt(f, DepthOfSearch => 3)
% numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true) -- FinalAttempt improves the estimate slightly

The computations performed when {\tt FinalAttempt} is set to {\tt true} are often slow, and often fail to improve the estimate, and for this reason, this option should be used sparingly. It is often more effective to increase the values of {\tt Attempts} or {\tt DepthOfSearch}, instead.

@@

i29 : f = -2*x^10*y^5 - x^5*y^9 - 2*x^3*y^10 + 2*x^2*y^8 - 2*x*y^9;

i30 : timing numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)

o30 = {.153461, .158333}
      -- 1.29156 seconds

o30 : Time

i31 : timing numeric fpt(f, DepthOfSearch => 3, Attempts => 4) -- a better 
      answer in less time

o31 = {.153333, .158333}
      -- .554321 seconds

o31 : Time

i32 : timing fpt(f, DepthOfSearch => 4) -- the exact answer in even less 
      time

       98
o32 = ---
      625
      -- .274657 seconds

o32 : Time
\end{MyVerbatim}
}
\medspace

% f = -2*x^10*y^5 - x^5*y^9 - 2*x^3*y^10 + 2*x^2*y^8 - 2*x*y^9;
% timing numeric fpt(f, DepthOfSearch => 3, FinalAttempt => true)
% timing numeric fpt(f, DepthOfSearch => 3, Attempts => 4) -- a better answer in less time
% timing fpt(f, DepthOfSearch => 4) -- the exact answer in even less time

As seen in several examples above, when the exact answer is not found, a list containing the endpoints of an interval containing the $F$-pure threshold of $f$ is returned. Whether that interval is open, closed, or a mixed interval depends on the options passed; if the option {\tt Verbose} is set to true, the precise interval will be printed.
{\small
\setstretch{.67}
\begin{MyVerbatim}

i33 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

i34 : fpt(f, DepthOfSearch => 3, FinalAttempt => true, Verbose => true)

@@

guessFPT narrowed the interval down to ( 3/40, 2/25 ) ...

@@

F-signature computation failed to find an improved lower bound ...

fpt failed to find the exact answer; try increasing the value of 
DepthOfSearch or Attempts.

fpt lies in the interval ( 3/40, 2/25 ).

        3   2
o34 = {--, --}
       40  25

o34 : List
\end{MyVerbatim}
}
\medspace

Finally, we point out that one can set the option {\tt IsLocal} from its default value of {\tt true} to {\tt false}, to compute the $F$-pure threshold globally. In other words, it computes the minimum of the $F$-pure threshold at all maximal ideals.

i1 : R = ZZ/7[x,y];

i2 : f = (y-1)^2-(x-1)^3;

i3 : fpt(f, IsLocal => false)

     5
o3 = -
     6

o3 : QQ

i4 : fpt(f)

o4 = infinity

o4 : InfiniteNumber
\end{MyVerbatim}
}
\medspace
% R = ZZ/7[x,y];
% f = (y-1)^2-(x-1)^3;
% fpt(f, IsLocal => false)
% fpt(f)

\noindent In this case, most options enabled by {\tt UseSpecialAlgorithms => true}
are ignored except for the check for simple normal crossings.  {\tt FinalCheck => true}
is also ignored.
